### 2024-02-02 paper.tex
very minor

To assess statistical significance of DMRs, we repeat the same procedure on permutations of the scBS data, i.e. the same data set with randomly shuffled cell labels.
The DMR t statistics obtained from permuted data are then used to estimate the false-discovery rate (FDR), yielding an adjusted p-value for each DMR.

@@

\texttt{scbs prepare} also computes a number of summary statistics for each cell, including the mean genome-wide methylation level and the number of observed CpG sites, i.e. the number of CpG sites that have sequencing coverage.
These summary statistics can be used to detect cells with poor quality.
The quality of each single cell methylome can furthermore be inspected with \texttt{scbs profile}, which computes the average methylation profile of a set of user-defined genomic regions such as transcription start sites (TSS) at single-cell resolution.
The TSS profile is a useful quality control plot since methylation shows a characteristic dip roughly \textpm1~kb around the transcription start site in mammalian genomes.
Cells that do not show this pattern, or cells with few observed CpG sites, may then be discarded with \texttt{scbs filter}.

@@

We also presented and published an open source software tool, called \texttt{scbs}, that provides an easy-to-use implementation of the described algorithms.
\new{It can start directly from the output of methylation callers such as Bismark \citep{bismark}, Biscuit \citep{biscuit} and methylpy \citep{methylpy} and produce a cell\texttimes region matrix.
We suggest iterative PCA as an approach to map the count matrix to a reduced-dimensional space overcoming the abundance of missing values.
However, one can also use to the same end established tools based on matrix factorisation such as MOFA+, \citep[][included in our benchmarks]{argelaguet2020mofa} or LIGER \citep{welch2019single}.
Once a low dimensional embedding is obtained, one can switch to well established tools from scRNA-seq analysis, including visualization tools such as t-SNE and UMAP, Leiden/Louvain clustering, pseudotime trajectory analyses, etc.
By offering these functions, \texttt{scbs} bridges a gap in the chain of existing tools that so far hindered practitioners in their data analysis.
Our implementation can handle datasets of various sizes up to a hundred thousand cells (see Supplementary Fig.~\ref{fig:umaps}B).}

@@

It may seem that we could use the standard procedure for the Welch t-test here, i.e. use the Welch-Satterthwaite formula to get an approximate degree of freedom and then calculate the tail probability of the corresponding t distribution.
However, this is unlikely to hold for two reasons:
First, the Welch-Satterthwaite degrees of freedom are only based on the number of cells per group and do not account for the fact that the read coverage might vary from cell to cell.
Second, the fusing of the DMRs obtained in the scanning step to obtain fused candidate DMRs would invalidate subsequent p-value based adjustment for multiple testing.

>>>>

To assess statistical significance of DMRs, we repeat the same procedure on permutations of the scBS data, i.e.\ the same data set with randomly shuffled cell labels.
The DMR t statistics obtained from permuted data are then used to estimate the false-discovery rate (FDR), yielding an adjusted p-value for each DMR.

@@

\texttt{scbs prepare} also computes a number of summary statistics for each cell, including the mean genome-wide methylation level and the number of observed CpG sites, i.e.\ the number of CpG sites that have sequencing coverage.
These summary statistics can be used to detect cells with poor quality.
The quality of each single cell methylome can furthermore be inspected with \texttt{scbs profile}, which computes the average methylation profile of a set of user-defined genomic regions such as transcription start sites (TSS) at single-cell resolution.
The TSS profile is a useful quality control plot since methylation shows a characteristic dip roughly \textpm1~kb around the transcription start site in mammalian genomes.
Cells that do not show this pattern, or cells with few observed CpG sites, may then be discarded with \texttt{scbs filter}.

@@

We also presented and published an open source software tool, called \texttt{scbs}, that provides an easy-to-use implementation of the described algorithms.
\new{It can start directly from the output of methylation callers such as Bismark \citep{bismark}, Biscuit \citep{biscuit} and methylpy \citep{methylpy} and produce a cell\texttimes region matrix.
We suggest iterative PCA as an approach to map the count matrix to a reduced-dimensional space overcoming the abundance of missing values.
However, one can also use to the same end established tools based on matrix factorisation such as MOFA+, \citep[][included in our benchmarks]{argelaguet2020mofa} or LIGER \citep{welch2019single}.
Once a low dimensional embedding is obtained, one can switch to well established tools from scRNA-seq analysis, including visualization tools such as t-SNE and UMAP, Leiden/Louvain clustering, pseudotime trajectory analyses, etc.
By offering these functions, \texttt{scbs} bridges a gap in the chain of existing tools that so far hindered practitioners in their data analysis.
Our implementation can handle datasets of various sizes up to a hundred thousand cells (see Fig.~\ref{fig:umaps}B).}

@@

It may seem that we could use the standard procedure for the Welch t-test here, i.e.\ use the Welch-Satterthwaite formula to get an approximate degree of freedom and then calculate the tail probability of the corresponding t distribution.
However, this is unlikely to hold for two reasons:
First, the Welch-Satterthwaite degrees of freedom are only based on the number of cells per group and do not account for the fact that the read coverage might vary from cell to cell.
Second, the fusing of the DMRs obtained in the scanning step to obtain fused candidate DMRs would invalidate subsequent p-value based adjustment for multiple testing.
